Adaptive group-regularized logistic elastic net regression

### 2019-03-12 docs/manuscript.Rnw
gren.R small bug fix

	In omics research, sources of such prior knowledge on the variables are often
	available. The information can, for example, come in the form of (a) results 
	on the same molecular features obtained in a previous study (e.g. $p$-values),
	(b) information from a publicly available database that summarizes the prior
	information on the molecular features involved (e.g., the Cancer Gene Census
	\cite[]{futreal_census_2004}), and (c) omics annotation (e.g. the location of 
	a gene on the chromosome). Although this information can rarely be directly
	included in the statistical analysis, it may still be useful and informative
	for the study at hand.

@@

	Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in
	spirit. Nevertheless, one clear advantage of our method over VBVS is the
	possibility of adaptive inclusion of external information on the features to
	aid in prediction and feature selection. Bayesian versions of support vector 
	machines have been used in classification problems as well 
	\cite[]{chakraborty_bayesian_2011}, but just as VBVS, these methods lack 
	adaptive inclusion of external information on the features. We emphasize that 
	our method is different from recent developments in grouped penalized 
	regression such as \cite{dondelinger_joint_nodate}, where instead of the 
	features, the observations are partitioned.

	In line with the above, our proposed method is (a) subjective: the 
	hyper-parameters are estimated from the data and not chosen on objective 
	grounds, (b) adaptive: prior information is automatically and subjectively 
	weighed with respect to its informativeness, (c) fast compared to full 
	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
	user: only the data and grouping of the features is required.

	The rest of the paper is structured as follows. We introduce the Bayesian 
	logistic generalised elastic net model in Section
	\ref{sec:model}. In Section \ref{seq:estimation} we derive a variational
	approximation to this model and use this novel approximation in the empirical
	Bayes estimation of multiple, group-specific penalty parameters. In Section
	\ref{sec:simulations} we compare the method in a simulation study. In Sections
	\ref{sec:colorectal} and \ref{sec:lymphnode}, we demonstrate the approach in 
	the two previously introduced data sets. In Section \ref{sec:discussion} 
	we conclude by both discussing some differences and
	parallels between the proposed method, termed \texttt{gren}, and other methods
	in the literature.

@@

	Assume we have a partitioning of the features in $G$ groups, such that each
	feature belongs to one group. Let $\G(g)$ be the feature index set of group 
	$g$ for $g=1, \dots, G$ and let $w_g \in \mathbb{R}_{>0}$ denote a 
	group-specific weight. In generalised elastic net regression, the penalised 
	likelihood is maximised to yield parameter estimates:
	\begin{align}
	  \hat{\bbeta} & := \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) - 
	  \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - 
	  \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 
	  \label{eq:weightedlik}\\
	  & = \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) -  \frac{\lambda_1}{2} 
	  \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - 
	  \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, 
	  \nonumber
	\end{align}
	where $\L(\y ; \bbeta)$ denotes the likelihood function of $\bbeta$ given the 
	observed data $\y = \begin{bmatrix} 
	y_1 & \cdots & y_n \end{bmatrix} \tr$, and 
	$\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ are the `global' penalty 
	parameters. We wrote $w^2_g = \lambda'_g$ to emphasise that these 
	group-specific weights may be interpreted as penalty multipliers. Note that
	the regular elastic net is recovered by setting $\forall g: \lambda'_g=1$.

@@

	The maximiser of the penalised likelihood in the elastic net, corresponds 
	to the posterior mode of a 
	Bayesian elastic model \cite[]{li_bayesian_2010,zou_regularization_2005}.
	\cite{li_bayesian_2010} show that the elastic net prior 
	(see SM Section [! section reference here] for details on this prior) may be 
	written as
	a computationally more convenient scale mixture of normals, with mixing 
	parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} 
	\tr$. Using this result, we write the generalised elastic net model in its 
	Bayesian form as:
	\begin{subequations}\label{eq:enmodel2}
		\begin{align}
		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i 
		\bbeta ) \right),  \\
		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0
		,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2
		}{\lambda_1^2}, \left(1,\infty \right) \right).
		\end{align}
	\end{subequations}
	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma 
	distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this
	Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} 
	\lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ 
	play the role of the hyperparameters in a Bayesian hierarchical model.

@@

	\subsection{Feature selection}
	Feature selection is often desirable in high-dimensional prediction problems 
	and omics data problems are no exception. For example, a selection of 
	biomarkers may lead 
	to a large cost reduction by supporting targeted assays. Bayesian feature
	selection is often done by inspection of posterior credible intervals. 
	Similarly as in frequentist hypothesis testing, we may select a feature if 
	zero is not contained in the credible interval. However, the Bayesian lasso's
	credible intervals (a special case of the elastic net) are known to suffer 
	from low frequentist coverage in sparse settings 
	\cite[]{castillo_bayesian_2015}. We often assume sparsity of the features in 
	omics research, so selection by credible intervals is generally inconsistent. 
	We therefore propose to select features in the frequentist paradigm.

@@

	\section{Simulations}\label{sec:simulations}
	To assess variable selection and predictive performance we conducted a 
	simulation study in which we compare \texttt{gren} to the regular elastic 
	net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016}, 
	sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, 
	group mimimax concave penalty (\texttt{grMCP}) 
	\cite[]{breheny_penalized_2009}, and the group exponential lasso 
	(\texttt{gel}) \cite[]{breheny_group_2015}. \texttt{GRridge} is similar to 
	\texttt{gren} in the sense that it estimates group-specific penalty 
	multipliers. The two main differences with \texttt{gren} are (i) the absence 
	of an $L_1$-norm penalty and (ii) the estimation procedure. The other methods 
	are extensions of the group lasso \cite[]{meier_group_2008, yuan_model_2006} 
	and not adaptive on the group level. In contrast to the original group lasso,
	all of them are able to select single features, instead of just on the group 
	level.

	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, 
	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
	and the elastic net, $\alpha$ determines the proportion of $L_1$-norm penalty,
	such that the first setting closely resembles the ridge setting, 
	where $\alpha=0$, while the third setting is similar to the lasso with 
	$\alpha=1$. In \texttt{SGL}, \texttt{grMCP}, and \texttt{gel}, $\alpha$
	determines the proportion of group-level penalization ($\alpha=0$) compared to
	$L_1$-norm penalty on the feature level ($\alpha=1$). For the methods that do 
	automatic variable selection, we estimate 
	the models for a range of model sizes.

	We simulated data according to five different scenarios:
	\begin{enumerate}
		\item differential signal between the groups and uniformly distributed model
		parameters,
		\item a large number of small groups of features,
		\item no differential signal between the groups, but strong correlations 
		within groups of features,
		\item differential signal between the groups and heavy-tailed distributed 
		model parameters, and
		\item a very sparse setting with no signal in some of the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section [! ref to section] of 
	the SM. In the first scenario, the distribution of the model parameters is far
	off from the assumed ridge/elastic net/lasso distributions, so we assume that 
	all methods will underperform. In the second scenario, we expect \texttt{gren}
	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
	estimate. The rest of the methods are conjectured to perform better, due to a
	lack of overfitting. In scenario three, we expect all methods to perform badly
	due to the high correlation between the features. The \texttt{gren} fourth 
	scenario is conjectured to give a better \texttt{gren} performance, due to the
	correctly assumed heavier-tailed distribution and differential signal between
	the groups. Scenario five is expected to lead to a bad performance in all 
	models, because of the sparseness of the truth and lack of differential 
	signal.

	In all scenarios we created a training set of $n=100$ samples to estimate the 
	models and a test set of $n_{\text{test}}=1000$ samples on which we compute 
	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
	from a $p$-dimensional zero-centred Gaussian distribution, where we set 
	$p=1000$. We let the covariance matrix of the features vary over the 
	scenarios. Likewise for the model parameters $\beta_j$, $j=1, \dots, p$. We 
	simulate the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
	of random variation in the simulations, we repeated every scenario 100 times 
	and report estimated penalty parameters and predictive performance measures 
	area under the receiver operator curve (AUC) and Brier skill score (BSS).

@@

	Additional results are presented in Section [! section ref here] of the SM.

	\section{Application to microRNAs in colorectal cancer}
	\label{sec:colorectal}
  As shortly touched upon in Section \ref{sec:introduction}, the data is from a 
  deep sequencing analysis on microRNAs 
  \cite[]{neerincx_combination_2018}. This study in 88 treated 
  colorectal cancer patients was conducted with the aim of predicting treatment 
  response, coded as either non-progressive/remission (70 patients) or 
  progressive 
  (18 patients). After pre-processing and normalisation, 2114 microRNAs 
  remained. In addition to the 2114 microRNAs, four clinical unpenalized 
  covariates were included in the analysis: prior use of adjuvant therapy 
  (binary), the type of systemic treatment regimen (ternary), age, and primary 
  tumor differentiation (binary).

  In a preliminary experiment on different subjects, the microRNA expression 
  levels of metastatic colorectal tumour tissue were compared to normal 
  non-colorectal tissue and primary colorectal tumour tissue was compared to
  primary colorectal tumour tissue \cite[]{neerincx_mir_2015}. This yielded 127
  microRNAs that were differentially expressed in both comparisons with
  $\text{FDR} \leq 0.001$, 95 with and $\text{FDR} \leq 0.05$, and 1893 not 
  differentially expressed microRNAs. We expect that incorporation of this 
  partitioning enhances therapy response 
  classification, because tumor-specific miRNAs might be more relevant than 
  non-specific ones.

  We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
  elastic net, and to the best performing
  group lasso extension, the composite minimax
  concave penalty (\texttt{cMCP}) by \cite{breheny_penalized_2009}. 
  For the methods with an extra tuning parameter 
  $\alpha$, we picked the one (or two) that gave the best performance and refer
  the reader to SM Section [! section ref here] for more comparisons with 
  different $\alpha$s and comparisons to the sparse
  group lasso by \cite{simon_sparse-group_2013} and group
  exponential lasso \cite[]{breheny_group_2015}.

  To estimate performance, we split the data into 61 training and 27 test 
  instances, stratified by treatment response. We estimated the models on the 
  training data and calculated AUC on the test data. We present AUC for a range 
  of model sizes, together with the estimated 
  penalty multipliers for \texttt{gren} and \texttt{GRridge} in Figure
  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}. More 
  performance measures are presented in Section [! section ref here]
  of the SM. In addition, we investigated the sensitivity of
  the multiplier estimation using in SM Section 
  [! section ref here]. We experienced numerical issues with 
  \texttt{GRridge} when cross-validating 
  the global penalty parameter on the training data. We therefore estimated this
  global parameter on the full data for \texttt{GRridge}.

  The penalty multipliers estimated by 
  \texttt{gren} and \texttt{GRridge} are according to expectation: the highly
  expressed group receives the lowest penalty, followed by the medium expressed
  group, while the non-expressed group receives the strongest penalty 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}a). 
  \texttt{gren} seems to outperform the other methods for a range of model 
  sizes in terms of AUC 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}b). 
  This is especially pronounced for the larger models. The performance of
  \texttt{cMCP} is somewhat unstable. We noticed that this is in general the 
  case for the group-lasso and its extensions (see SM [! section ref here]). 
  Overall, it seems as though the inclusion of the extra information on the 
  features benefited predictive performance: both \texttt{GRridge} and 
  \texttt{gren} outperform their respective non-group-regularized versions, 
  albeit it only slightly for \texttt{GRridge}.
<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Results for the colorectal cancer example with (a) the estimated penalty multipliers and (b) estimated AUC.", out.width="100%", fig.asp=1/2>>=
@

  \section{Application to RNAseq in lymph node cancer}\label{sec:lymphnode}
  The aim of this study is to predict lymph node metastasis in patients 
  with HPV-negative oral cancer using sequenced RNAs from TCGA 
  \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
  3096 transformed and normalised TCGA RNASeqv2 profiles of head-and-neck 
  squamous cell carcinomas on 133 patients with HPV-negative tumours in the oral
  cavity. Of these patients, 76 suffered from lymph node metastasis, while 57 
  did not. For a thorough introduction of these data, see 
  \cite{te_beest_improved_2017}.

  As external feature information, we considered two sources: (a) the 
  cis-correlation between the RNASeqv2 data and TCGA DNA copy numbers on 
  the same patients, 
  quantified by Kendall's $\tau$ and binned into five groups. We also considered
  (b) $p$-values from an external microarray data set described in 
  \cite{mes_prognostic_2017}, again binned into five groups. We expect features 
  with a large positive Kendall's $\tau$ to be more important in 
  metastasis prediction \cite[]{masayesva_gene_2004}. Likewise, we expect 
  features with low $p$-values to be more important.

  We compared \texttt{gren} to the same methods as in Section 
  \ref{sec:colorectal}. However, because the co-data consists of two partitions 
  with overlapping groups, we use an extension of \texttt{cMCP} that allows for
  overlapping groups \cite[]{zeng_overlapping_2016}.
  We estimated AUC for the methods on an independent validation set of 97 
  HPV-negative oral cancer patients \cite[]{mes_prognostic_2017}. The features 
  in this validation set were
  microarrays, normalised to account for a different scale of the transformed 
  RNAseq data. We present the results on this validation set in Figure 
  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. Other 
  comparisons and performance measures are presented in the SM Section 
  [! section ref here].

  The estimated penalty multipliers follow the expected pattern 
  (Figure \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}a): 
  larger cis-correlations receive smaller penalties. 
  Likewise, smaller $p$-values are penalized less. This pattern is slightly 
  less pronounced in \texttt{gren} with $\alpha=0.95$ compared to the other 
  methods. The predictive performance of
  \texttt{gren} is slighty better than the other methods for a range of model
  sizes. We also note that the overlapping group \texttt{cMCP} is much more
  stable than the regular \texttt{cMCP} (see in SM [! section ref here]), 
  where we create one paritioning by cross-tabulation of the two paritions. 
  In this example, ridge and \texttt{GRridge} perform almost identical.
<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Results for the lymph node metastasis example with (a) the estimated penalty multipliers for the cis-correlations, (b) the estimated penalty multipliers for the p-values, and (c) the estimated AUC.", out.width="100%", fig.asp=1>>=
@

>>>>

	In omics research, sources of such prior knowledge on the variables are often
	available. The information can, for example, come in the form of (a) results 
	on the same molecular features obtained in a previous study (e.g. $p$-values),
	(b) information from a publicly available database that summarizes the prior
	information on the molecular features involved (e.g., the Cancer Gene Census
	\cite[]{futreal_census_2004}), and (c) omics annotation (e.g. a gene's 
	chromosome location). Although this information can rarely be directly
	included in the statistical analysis, it may still be useful and informative
	for the study at hand.

@@

	Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in
	spirit. Nevertheless, one clear advantage of our method over VBVS is the
	possibility of adaptive inclusion of external information on the features to
	aid in prediction and feature selection. Bayesian versions of support vector 
	machines have been used in classification problems as well 
	\cite[]{chakraborty_bayesian_2011}, but these methods also lack 
	adaptive inclusion of external information on the features.

	In line with the above, our proposed method is (a) subjective: the 
	hyper-parameters are estimated from the data and not chosen on objective 
	grounds, (b) adaptive: prior information is automatically and subjectively 
	weighed with respect to its informativeness, (c) fast compared to full 
	Bayesian analysis or cross-validation, and (d) requires minimal input by the 
	user: only the data and grouping of the features is required. We emphasize that 
	our method is different from recent developments in grouped penalized 
	regression such as \cite{dondelinger_joint_nodate}, where instead of the 
	features, the observations are partitioned.

	The rest of the paper is structured as follows. We introduce the Bayesian 
	logistic generalised elastic net model in Section
	\ref{sec:model}. In Section \ref{seq:estimation} we derive a variational
	approximation to this model and use this novel approximation in the empirical
	Bayes estimation of multiple, group-specific penalty parameters. In Section
	\ref{sec:simulations} we compare the method in a simulation study. In Sections
	\ref{sec:colorectal} and \ref{sec:lymphnode}, we demonstrate the approach in 
	two data sets. In Section \ref{sec:discussion} 
	we conclude by both discussing some differences and
	parallels between the proposed method, termed \texttt{gren}, and other methods
	in the literature.

@@

	Assume we have a partitioning of the features in $G$ groups, such that each
	feature belongs to one group. Let $\G(g)$ be the feature index set of group 
	$g$ for $g=1, \dots, G$ and let $w_g \in \mathbb{R}_{>0}$ denote a 
	group-specific weight. In a generalised elastic net regression, the penalised 
	likelihood is maximised to yield parameter estimates:
	\begin{align}
	  \hat{\bbeta} & := \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) - 
	  \frac{\lambda_1}{2} \sum_{g=1}^G \sum_{j \in \G(g)} |w_g \cdot \beta_j| - 
	  \frac{\lambda_2}{2} \sum_{g=1}^G \sum_{j \in \G(g)}\( w_g \cdot \beta_j\)^2 
	  \label{eq:weightedlik}\\
	  & = \underset{\bbeta}{\argmax} \log \L(\y ; \bbeta) -  \frac{\lambda_1}{2} 
	  \sum_{g=1}^G \sqrt{\lambda'_g} \sum_{j \in \G(g)} | \beta_j| - 
	  \frac{\lambda_2}{2} \sum_{g=1}^G \lambda'_g \sum_{j \in \G(g)} \beta_j^2, 
	  \nonumber
	\end{align}
	where $\L(\y ; \bbeta)$ denotes the likelihood function of $\bbeta$ given the 
	observed data $\y = \begin{bmatrix} 
	y_1 & \cdots & y_n \end{bmatrix} \tr$, and 
	$\lambda_1, \lambda_2 \in \mathbb{R}_{>0}$ are the `global' penalty 
	parameters. We wrote $w^2_g = \lambda'_g$ to emphasise that these 
	group-specific weights may be interpreted as penalty multipliers. Note that
	the regular elastic net is recovered by setting $\forall g: \lambda'_g=1$.

@@

	The maximiser of the penalised likelihood in the elastic net, corresponds 
	to the posterior mode of a 
	Bayesian elastic model \cite[]{li_bayesian_2010,zou_regularization_2005}.
	\cite{li_bayesian_2010} show that the elastic net prior 
	(see Supplementary Material (SM) Section [! section reference here] for 
	details on this prior) may be written as
	a computationally more convenient scale mixture of normals, with mixing 
	parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} 
	\tr$. Using this result, we write the generalised elastic net model in its 
	Bayesian form as:
	\begin{subequations}\label{eq:enmodel2}
		\begin{align}
		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i 
		\bbeta ) \right),  \\
		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0
		,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2
		}{\lambda_1^2}, \left(1,\infty \right) \right).
		\end{align}
	\end{subequations}
	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma 
	distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this
	Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} 
	\lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ 
	play the role of the hyperparameters in a Bayesian hierarchical model.

@@

	\subsection{Feature selection}
	Feature selection is often desirable in high-dimensional prediction problems 
	and omics data problems are no exception. For example, biomarker selection 
	may lead 
	to a large reduction in costs by supporting targeted assays. Bayesian feature
	selection is often done by inspection of posterior credible intervals. 
	Similarly as in frequentist hypothesis testing, we may select a feature if 
	zero is not in the credible interval. However, the Bayesian lasso's 
	(a special case of the elastic net) credible intervals  are known to suffer 
	from low frequentist coverage in sparse settings 
	\cite[]{castillo_bayesian_2015}. We often assume sparsity of the features in 
	omics research, so selection by credible intervals is generally inconsistent. 
	We therefore propose to select features in the frequentist paradigm.

@@

	\section{Simulations}\label{sec:simulations}
	To assess variable selection and predictive performance we conducted a 
	simulation study in which we compared \texttt{gren} to the regular elastic 
	net and ridge models, \texttt{GRridge} \cite[]{van_de_wiel_better_2016}, 
	sparse group lasso (\texttt{SGL}) \cite[]{simon_sparse-group_2013}, 
	composite mimimax concave penalty (\texttt{cMCP}) 
	\cite[]{breheny_penalized_2009}, and the group exponential lasso 
	(\texttt{gel}) \cite[]{breheny_group_2015}. \texttt{GRridge} is similar to 
	\texttt{gren} in the sense that it estimates group-specific penalty 
	multipliers. The two main differences with \texttt{gren} are (i) the absence 
	of an $L_1$-norm penalty and (ii) the estimation procedure. The other methods 
	are extensions of the group lasso and not adaptive on the group level. 
	However, in contrast to the original group lasso,
	they select single features, instead of just whole groups.

	For the regular elastic net, \texttt{gren}, \texttt{SGL}, \texttt{cMCP}, 
	and \texttt{gel}, we fix $\alpha \in \{0.05, 0.5, 0.95 \}$. In \texttt{gren}
	and the elastic net, $\alpha$ determines the proportion of $L_1$-norm penalty,
	such that the first setting closely resembles the ridge setting, 
	where $\alpha=0$, while the third setting is similar to the lasso with 
	$\alpha=1$. In \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}, $\alpha$
	determines the ratio of group-level penalization ($\alpha=0$) to 
	feature level ($\alpha=1$) penalization. For the methods that do 
	automatic feature selection, we estimate 
	the models for a range of model sizes.

	We simulated data according to five different scenarios:
	\begin{enumerate}
		\item differential signal between the groups and uniformly distributed model
		parameters,
		\item a large number of small groups of features,
		\item no differential signal between the groups, but strong correlations 
		within groups of features,
		\item differential signal between the groups and heavy-tailed distributed 
		model parameters, and
		\item a very sparse setting with no signal in some of the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section [! section ref here] 
	of 
	the SM. In the first scenario, the distribution of model parameters is far
	from the assumed ridge/elastic net/lasso distributions, so we expect
	all methods to underperform. In the second scenario, we expect \texttt{gren}
	and \texttt{GRridge} to overfit due to the large number of hyperparameters to 
	estimate, while we expect the other methods to perform better, due to a
	lack of overfitting. In scenario three, we expect all methods to underperform
	due to the high correlations between the features. The fourth 
	scenario is conjectured to yield a better \texttt{gren} performance, since
	\texttt{gren} assumes a heavier-tailed distribution and is designed to 
	incorporate a differential signal between
	the groups. Scenario five is expected to yield a bad performance for all 
	methods, because of the sparse truth and lack of differential 
	signal.

	In all scenarios we created a training set of $n=100$ samples to estimate the 
	models and a test set of $n_{\text{test}}=1000$ samples to compute 
	the performance measures. The features $\x_i$, $i=1, \dots, n$, are taken 
	from a $p$-dimensional zero-centred Gaussian distribution, where we set 
	$p=1000$. We varied the covariance matrix of the features and simulation of 
	the model parameters $\beta_j$, $j=1, \dots, p$ between the
	scenarios (see SM Section [! section ref here] for the exact 
	parametrisations). We 
	simulated the outcomes $y_i$, $i=1, \dots, n$, from a binary logistic model: 
	$y_i \sim \mathcal{B}(1, \expit (\x_i \tr \bbeta))$. To mitigate the influence
	of random variation, we repeated each scenario 100 times 
	and report estimated penalty parameters and predictive performance measures 
	area under the receiver operator curve (AUC) and Brier skill score (BSS) in 
	Table [! table ref here].

	- description of results

@@

	Additional results and performance measures are presented in Section 
	[! section ref here] of the SM.

	\section{Application to microRNAs in colorectal cancer}
	\label{sec:colorectal}
  We investigated the performance of \texttt{gren} on data from a 
  deep sequencing study in microRNAs 
  \cite[]{neerincx_combination_2018}. The aim of the study was to predict 
  treatment response in 88 treated colorectal cancer patients, coded as either
  non-progressive/remission (70 patients) or progressive 
  (18 patients). After pre-processing and normalisation, 2114 microRNAs 
  remained. In addition to the microRNAs, four unpenalized clinical
  covariates were included in the analysis: prior use of adjuvant therapy 
  (binary), the type of systemic treatment regimen (ternary), age, and primary 
  tumor differentiation (binary).

  In a preliminary experiment on different subjects, the microRNA expression 
  levels of metastatic colorectal tumour tissue were compared to normal 
  non-colorectal tissue and primary colorectal tumour tissue was compared to
  primary colorectal tumour tissue \cite[]{neerincx_mir_2015}. This yielded 127
  microRNAs that were differentially expressed in both comparisons with
  $\text{FDR} \leq 0.001$, 95 with and $\text{FDR} \leq 0.05$, and 1893 
  microRNAs that were not differentially expressed. We expect that incorporation
  of this partitioning enhances therapy response 
  classification, because tumor-specific microRNAs are likely to be more 
  relevant than non-specific microRNAs.

  We compared the performance of \texttt{gren} to ridge, \texttt{GRridge}, 
  elastic net, \texttt{SGL}, \texttt{cMCP}, and \texttt{gel}. Of the latter 
  three methods, we only present \texttt{cMCP} here; the results for 
  \texttt{SGL} and \texttt{gel} are presented in the SM Section 
  [! section ref here]. For the methods with a
  tuning parameter $\alpha$, we picked the $\alpha$('s) that gave the best
  performance and refer
  the reader to SM Section [! section ref here] for comparisons with 
  different $\alpha$'s.

  To estimate performance, we split the data into 61 training and 27 test 
  instances, stratified by treatment response. We estimated the models on the 
  training data and calculated AUC on the test data. We present the AUC for a
  range of model sizes, together with the estimated 
  penalty multipliers for \texttt{gren} and \texttt{GRridge} in Figure
  \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}. More 
  performance measures are presented in Section [! section ref here]
  of the SM. In addition, we investigated the sensitivity of
  the multiplier estimation on this dataset in SM Section 
  [! section ref here]. We experienced numerical issues with 
  \texttt{GRridge} when cross-validating 
  the global $\lambda$ on the training data. We therefore estimated this
  global parameter on the full data for \texttt{GRridge}.

  The penalty multipliers estimated by 
  \texttt{gren} and \texttt{GRridge} are according to expectation: the highly
  expressed group receives the lowest penalty, followed by the medium expressed
  group, while the non-expressed group receives the strongest penalty 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}a). 
  \texttt{gren} seems to outperform the other methods for a range of model 
  sizes in terms of AUC 
  (Figure \ref{fig:barplot_lines_micrornaseq_colorectal_cancer_res1_auc}b). 
  This is especially pronounced for the larger models. The performance of
  \texttt{cMCP} is somewhat unstable. This unstable estimation is an issue for   
  all group lasso extensions that we investigated.
  (SM [! section ref here]). Overall, it seems as though the inclusion of the
  extra information on the 
  features benefited predictive performance: both \texttt{GRridge} and 
  \texttt{gren} outperform their respective non-group-regularized versions, 
  albeit it only slightly for \texttt{GRridge}.
<<barplot_lines_micrornaseq_colorectal_cancer_res1_auc, fig.cap="Estimated (a) penalty multipliers and (b) AUC in the colorectal cancer example.", out.width="100%", fig.asp=1/2>>=
@

  \section{Application to RNAseq in lymph node cancer}\label{sec:lymphnode}
  The aim of our second example study was to predict lymph node metastasis in
  oral cancer patients using sequenced RNAs from TCGA 
  \cite[]{the_cancer_genome_atlas_network_comprehensive_2015}. The features are 
  3096 transformed and normalised TCGA RNASeqv2 profiles of head-and-neck 
  squamous cell carcinomas on 133 patients with HPV-negative tumours in the oral
  cavity. Of these patients, 76 suffered from lymph node metastasis, while 57 
  did not. For a thorough introduction of these data, see 
  \cite{te_beest_improved_2017}.

  We considered two sources of external feature information: (a) the 
  cis-correlations between the RNASeqv2 data and TCGA DNA copy numbers on 
  the same patients, quantified by Kendall's $\tau$ and binned into five groups.
  In addition, we used (b) $p$-values from a microarray data set
  described in \cite{mes_prognostic_2017}, again binned into five groups. 
  We expected features 
  with a large positive Kendall's $\tau$ to be more important in 
  metastasis prediction \cite[]{masayesva_gene_2004}. Likewise, we expect 
  features with low external $p$-values to be more important.

  We compared \texttt{gren} to the same methods as in Section 
  \ref{sec:colorectal}. However, since the feature information
  consists of two partitions with overlapping groups, we used an extension of
  \texttt{cMCP} that allows for
  overlapping groups \cite[]{zeng_overlapping_2016}.
  We estimated AUC for the methods on an independent validation set of 97 
  patients \cite[]{mes_prognostic_2017}. The features 
  in this validation set were
  microarray expression levels, normalised to account
  for a difference in scale compared to the RNAseq training data. We present the
  results on this validation set, together with the estimated penalty 
  multipliers for \texttt{gren} and \texttt{GRridge}, in Figure 
  \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}. Additional
  comparisons and performance measures are presented in the SM Section 
  [! section ref here].

  The estimated penalty multipliers follow the expected pattern 
  (Figure \ref{fig:barplot_lines_rnaseq_lymph_node_metastasis_res1_auc}a): 
  larger cis-correlations receive smaller penalties. 
  Likewise, smaller external $p$-values are penalized less. This pattern is
  slightly less pronounced in \texttt{gren} with $\alpha=0.95$ compared to the
  other methods. The predictive performance of
  \texttt{gren} is slighty better than the other methods for a range of model
  sizes. In this example, ridge and \texttt{GRridge} perform almost 
  identical, while \texttt{gren} outperforms the regular elastic net.
<<barplot_lines_rnaseq_lymph_node_metastasis_res1_auc, fig.cap="Estimated (a) penalty multipliers for the cis-correlations and (b) external $p$-values, and (c) estimated AUC in the lymph node metastasis example.", out.width="100%", fig.asp=1>>=
@

	\section{Discussion}\label{sec:discussion}
	In a taxonomy of Bayesian methods, \texttt{gren} may be considered a 
	local shrinkage model, as opposed to the global-local shrinkage priors that 
	\cite{polson_local_2012,bernardo_shrink_2011} discuss. They characterise 
	certain desirable properties of these global-local shrinkage priors in high 
	dimensions, which, for example, the horseshoe possesses 
	\cite[]{carvalho_horseshoe_2010,carvalho_handling_2009}. In our case, global 
	shrinkage would imply adding another hyperprior for the global $\lambda_1$ 
	and $\lambda_2$ (or $\alpha$ and $\lambda$) hyperparameters. We argue however,
	that if the groups are informative, the empirical Bayes estimation of the 
	(semi-)global shrinkage parameters $\lambda'_g$ may be more beneficial than 
	full Bayes shrinkage of the global penalty parameters, because the latter does
	not use any known structure to model the variability in the hyperparameters. 
	Nonetheless, an interesting direction of future research is the extension of 
	the group-regularized elastic net to a group-regularized horseshoe model, 
	since the horseshoe has been shown to handle sparsity well and render better
	coverage of credibility intervals than lasso-type priors 
	\cite[]{van_der_pas_horseshoe_2014}.

	Although our method can be considered weakly adaptive and is essentially 
	a reweighted elastic net, it is different from 
	the adaptive lasso \cite[]{van_de_geer_adaptive_2011,huang_adaptive_2008,
	zhang_adaptive_2007,zou_adaptive_2006} and adaptive elastic net 
	\cite[]{zou_adaptive_2009} in the sense that it adapts to external information
	rather than to the primary data. It also differs in the scale of adaptation: 
	in the adaptive lasso and elastic net the adaptive weights are feature 
	specific, while in our case they are estimated on the group level, rendering
	the adaptation more robust. As can be seen from both the simulations in 
	Section \ref{sec:simulations} and the applications to different outcomes 
	(prognostic or diagnostic) and marker types 
	(RNAseq, microRNAs, or metabolomics) in Sections \ref{sec:colorectal}--
	\ref{sec:lymphnode}, and SM Sections [! section refs here], adaptation to 
	external data may be beneficial for prediction and feature selection. 
	We believe that this is due to the `borrowing of information' effect: 
	estimates of feature effects that behave similarly are shrunken similarly, 
	yielding overall, better estimates.
